---
output: latex_fragment
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(comment = "")
options(width = 60, digits = 2)
library(reda)
suppressPackageStartupMessages(library(ggplot2))
data(readmission, package = "frailtypack")
df0 <- subset(readmission, !(id %in% c(60, 109, 280)))
```

The mean cumulative function (MCF) is often the focus in the nonparametric analysis of
recurrent events.
Let $M_i(t)=\mathbb{E}\{N_i(t)\}$ denote the MCF of
$N_i(t)$.  The Nelson-Aalen estimator \citep{nelson2003siam}
for the MCF takes the form
<!--exploring the trend of recurrent events has the form:-->
$$\widehat{M}(t) = \int_0^t \frac{dN(s)}{\delta(s)},$$
where $dN(s)=\sum_{i=1}^k dN_i(s)$, $\delta(s) =
\sum_{i=1}^k \delta_i(s)$, $dN_i(s)$ and $\delta_i(s)$ is, respectively, the
jump size and at-risk indicator of process $i$ at time $s$.
The MCF can be visualized by plotting the `Recur` object with argument
`mcf = TRUE` when the \pkg{reReg} package is active, e.g., `plot(obj, mcf = TRUE)`.
<!--The \pkg{reReg} package also offers the `plotMCF()` function for plotting MCFs
stratified by discrete variables. -->
Alternatively, the `mcf()` function from the \pkg{reda} package provides a more
sophisticated approach to plot MCFs and make inference.
The following example uses the `mcf()` function to visualize MCF estimates
stratified by whether the patients received chemotherapy.
<!-- We may use the `mcf()` function and the associated `plot()` method provided by
the \pkg{reda} package to visualize MCF estimates stratified by whether the
patients received chemotherapy as follows.
Terminal events are ignored here for ease of illustration. -->

```{r plot-sampleMcf-code, eval = FALSE}
re_mcf <- mcf(fn, data = df0)
plot(re_mcf, conf.int = TRUE, lty = 1:2) +
    ggplot2::theme(legend.position = "bottom") # Fig. 3
```

```{r plot-sampleMcf, echo = FALSE, fig.path = 'images/', fig.show = 'hide', fig.width = 14, fig.height = 12}
re_mcf <- mcf(Recur(t.stop, id, event) ~ chemo, data = df0)
plot(re_mcf, conf.int = TRUE, lty = 1:2) +
    theme(legend.position = "bottom",
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8),
          legend.box.margin = margin(-12, -10, -10, -10))
```
\vspace*{-.3cm}
\begin{figure}[H]
    \centering
    \begin{minipage}{0.245\textwidth}
        \centering
        \includegraphics[scale = .125]{images/plot-sampleMcf-1}
        \caption{Stratified by \texttt{chemo}}
    \end{minipage}\hfill
    \begin{minipage}{0.245\textwidth}
        \centering
        \includegraphics[scale = .125]{images/mcfdiffplot-1}
        \caption{MCF difference}
    \end{minipage}
\end{figure}
<!--The `plot()` method return a `ggplot2` object \citep{hadley2016ggplot2} so that
users may further customize the plot easily.-->
The MCF difference between two groups can be tested
via `mcfDiff.test()`, which implements the two-sample pseudo-score tests of
\citep{cook1996biometrics}.
The following results indicate the MCF estimates are statistically different at
a significance level of 0.05.
The MCF difference can be plotted with directly by `plot(mcfDiff(re_mcf))`,
as shown in Fig. 4.
```{r mcfdiff, echo = TRUE}
mcfDiff.test(re_mcf)
```
```{r mcfdiffplot, echo = FALSE, fig.path = 'images/', fig.show = 'hide', fig.width = 14, fig.height = 12}
plot(mcfDiff(re_mcf))
```
